MATLAB: Read Me

### 2020-04-02 MATLAB_Files/MATLAB_readMe/MATLAB_readMe.tex
I am checking that the empirical moments match with the analytic ones.

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{SDE first moment (1/2):} \label{m1}
\alert{Check if still correct for the new model!!!}\\
Given some measurement $v_{t_{n-1}}$, we want to compute the first moment at time $t_n$. The exact first moment $m_1(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

@@

The code is automatically imported from the MATLAB script.

@@

\alert{Check if still correct for the new model!!!}\\
Given some measurement $v_{t_{n-1}}$, we want to compute the second moment at time $t_n$. The exact second moment $m_2(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

\begin{equation*}
\begin{cases}
\dif m_2(s)&=\alert{\left[-2(1+\alpha)m_2(s)\theta(s)+2\alpha\theta(s)m_1(s)(1-2p(s))+2\alpha\theta(s)p(s)(1-p(s))\right]\dif s,}\\
&=2\theta(s)\left[-(1+\alpha)m_2(s)+\alpha m_1(s)(1-2p(s))+\alpha p(s)(1-p(s))\right]\dif s,\\
m_2(t_{n-1})&=v^2_{t_{n-1}}.
\end{cases}
\end{equation*}
We compute a linear interpolation for the functions $\theta(s)$ and $p(s)$. After, we solve the ODE using Forward-Euler:
{\scriptsize
\begin{equation*}
m_2(s_n)=m_2(s_{n-1})+2\theta(s_{n-1})\left[-(1+\alpha)m_2(s_{n-1})+\alpha m_1(s_{n-1})(1-2p(s_{n-1}))+\alpha p(s_{n-1})(1-p(s_{n-1}))\right]\Delta s.
\end{equation*}}
We use the same discretization points for both $m_1(s)$ and $m_2(s)$.

@@

\begin{center}
\begin{tabular}{|c|}
\toprule
{\footnotesize
\lstinputlisting{../moment_2.m}
}\\
\bottomrule
\end{tabular}
\end{center}

@@

To ensure that the analytical solutions is always in $]0,1[$, we choose the drift parameter to be
\begin{equation*}
\theta(t) = \max\left(\theta_0,\frac{\theta_0\alpha+2|\dot{p}^\epsilon(t)|}{2\min(p^\epsilon(t),1-p^\epsilon(t))}\right),\quad\theta_0>0.
\end{equation*}
\quad\\
\quad\\
\begin{center}
\begin{tabular}{|c|}
\toprule
{\tiny
\lstinputlisting{../theta_t.m}
}\\
\bottomrule
\end{tabular}
\end{center}
{\tiny Recall that we normalized w.r.t. time $T$. The normalized $\hat{\theta}_t$ is $\hat{\theta}_t=\theta_tT$. Notice that in the definition of $\theta_t$, if we use $\hat{\theta}_0=\theta_0T$ and the derivative of $p_t$ w.r.t. $\dif \hat{t}$ instead of w.r.t. $\dif t$, then we automatically have $\hat{\theta}_t$. See slide \ref{Norm}.}

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=0.32\textwidth]{../Results/likelihood/Log-Likelihood.eps}
\includegraphics[width=0.32\textwidth]{../Results/likelihood/Log-Likelihood_refined.eps}
\includegraphics[width=0.32\textwidth]{../Results/likelihood/Log-Likelihood_more_refined.eps}
\caption{We use the code \textbf{plotLogLikelihood.m} to create this plots. We used about 18 thousand transitions, and we can also see the initial guess.}
\end{figure}

@@

\item {\color{blue} $Z_t=\frac{1}{\sqrt{2\alpha\theta_0}}\arcsin(2(V_t+p_t)-1)$.}
\item {\color{red} $Z_t=\sqrt{\frac{2}{\alpha\theta_0}}\arcsin(\sqrt{1-V_t-p_t})$.}

\end{itemize}
\quad\\
\quad\\
They have the same partial derivatives. The main difference appears in the resulting SDE for $Z_t$. In the {\color{blue} first case}, we have a $\sin(\sqrt{2\alpha\theta_0}Z_t)$ in the drift, while in the \alert{second case}, we have $\cos(\sqrt{2\alpha\theta_0}Z_t)$.\\
\quad\\
For now, we will use the {\color{blue}first candidate}.

@@

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Lamperti data:}

We are following the {\color{blue}first candidate}. Then, we caa

@@

The Lamperti SDE is:
\begin{align*}
\dif Z_t= \underbrace{\left[  \frac{(\alpha \theta_0 - \theta_t) {\color{blue}\sin(\sqrt{2 \alpha \theta_0 } Z_t) }- \theta_t (1 - 2 p_t) + 2  \dot{p}_t }{\sqrt{2 \alpha \theta_0} \cos{(\sqrt{2 \alpha \theta_0} Z_t)}}  \right]}_{\coloneqq b(Z_t)} \dif t + \dif W_t . 
\end{align*}
Here we follow Bayesian Filtering and Smoothing, Chapter 9. We use the linearization-based approximation for the moments.

@@

>>>>

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{SDE first moment (1/2):} \label{m1}
Given some measurement $v_{t_{n-1}}$, we want to compute the first moment at time $t_n$. The exact first moment $m_1(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

@@

@@

Given some measurement $v_{t_{n-1}}$, we want to compute the second moment at time $t_n$. The exact second moment $m_2(s)$ for $s\in[t_{n-1},t_n]$ is the solution of the ODE

\begin{equation*}
\begin{cases}
\dif m_2(s)&=\left[\alert{-2m_2(s)(\theta(s)+\alpha\theta_0)}+2\alpha\theta_0m_1(s)(1-2p(s))+{\color{blue}2\alpha\theta_0p(s)(1-p(s))}\right]\dif s,\\
m_2(t_{n-1})&=v^2_{t_{n-1}}.
\end{cases}
\end{equation*}
We compute a linear interpolation for the functions $\theta(s)$ and $p(s)$. After, we solve the ODE using Forward-Euler:
{\scriptsize
\begin{equation*}
m_2(s_n)=m_2(s_{n-1})+\left[\alert{-2m_2(s_{n-1})(\theta(s_{n-1})+\alpha\theta_0)}+2\alpha\theta_0m_1(s_{n-1})(1-2p(s_{n-1}))+{\color{blue}2\alpha\theta_0p(s_{n-1})(1-p(s_{n-1}))}\right]\Delta s.
\end{equation*}}
We use the same discretization points for both $m_1(s)$ and $m_2(s)$.

@@

\begin{center}
\begin{tabular}{|c|}
\toprule
{\tiny
\lstinputlisting{../moment_2.m}
}\\
\bottomrule
\end{tabular}
\end{center}

@@

To ensure that the analytical solutions is always in $]0,1[$, we choose the drift parameter to be
\begin{equation*}
\theta(t) = \max\left(\theta_0,\frac{\theta_0\alpha+|\dot{p}^\epsilon(t)|}{\min(p^\epsilon(t),1-p^\epsilon(t))}\right),\quad\theta_0>0.
\end{equation*}
\quad\\
\quad\\
\begin{center}
\begin{tabular}{|c|}
\toprule
{\tiny
\lstinputlisting{../theta_t.m}
}\\
\bottomrule
\end{tabular}
\end{center}
{\tiny Recall that we normalized w.r.t. time $T$. The normalized $\hat{\theta}_t$ is $\hat{\theta}_t=\theta_tT$. Notice that in the definition of $\theta_t$, if we use $\hat{\theta}_0=\theta_0T$ and the derivative of $p_t$ w.r.t. $\dif \hat{t}$ instead of w.r.t. $\dif t$, then we automatically have $\hat{\theta}_t$. See slide \ref{Norm}.}

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=0.32\textwidth]{../Results/likelihood/normal/Log-Likelihood.eps}
\includegraphics[width=0.32\textwidth]{../Results/likelihood/normal/Log-Likelihood_refined.eps}
\includegraphics[width=0.32\textwidth]{../Results/likelihood/normal/Log-Likelihood_more_refined.eps}
\caption{We use the code \textbf{plotLogLikelihood.m} to create this plots. We used about 18 thousand transitions, and we can also see the initial guess.}
\end{figure}

@@

\item {\color{blue} $Z_t=\frac{1}{\sqrt{2\alpha\theta_0}}\arcsin(2(V_t+p_t)-1)$.}
\item {\color{red} $Z_t=-\sqrt{\frac{2}{\alpha\theta_0}}\arcsin(\sqrt{1-V_t-p_t})$.}

\end{itemize}
\quad\\
\quad\\
They have the same partial derivatives. The main difference appears in the resulting SDE for $Z_t$.\\
\quad\\
For now, we will use the {\color{blue}first candidate}.

@@

@@

The Lamperti SDE is:
\begin{align*}
\dif Z_t= \underbrace{\left[  \frac{(\alpha \theta_0 - \theta_t) {\color{black}\sin(\sqrt{2 \alpha \theta_0 } Z_t) }- \theta_t (1 - 2 p_t) + 2  \dot{p}_t }{\sqrt{2 \alpha \theta_0} \cos{(\sqrt{2 \alpha \theta_0} Z_t)}}  \right]}_{\coloneqq b(Z_t)} \dif t + \dif W_t . 
\end{align*}
Here we follow Bayesian Filtering and Smoothing, Chapter 9. We use the linearization-based approximation for the moments.\\
\quad\\
As $X_t\in(\alert{0},1)$, we have that $Z_t\in\left(\alert{-\frac{\pi}{2}\frac{1}{\sqrt{2\alpha\theta_0}}},\frac{\pi}{2}\frac{1}{\sqrt{2\alpha\theta_0}}\right)$.

@@

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Lamperti Log-likelihood evaluation (1/2): CODE}

\begin{center}
\begin{tabular}{|c|}
\toprule
{\tiny
\lstinputlisting{../log_LH_evaluation_L.m}
}\\
\bottomrule
\end{tabular}
\end{center}

\setbeamercolor{background canvas}{bg=white!10}
\begin{frame}\frametitle{Lamperti Log-likelihood evaluation (2/2):}

\begin{figure}[ht!]
\centering
\includegraphics[width=0.5\textwidth]{../Results/likelihood/lamperti/Log-Likelihood.eps}
\caption{We use the code \textbf{plotLogLikelihood.m} to create this plots. We used about 18 thousand transitions.}
\end{figure}
